---
title: "Tests of Proportions"
author: "Brigid Wilson"
date: "September, 2014"
output: html_document
---

R functions can be used to calculate achieved power or required samples size for tests of proportions across indepdendent samples. 
```{r}
power.prop.test(n = 25, p1 = 0.28, p2 = 0.52, sig.level = 0.05,
                power = NULL,
                alternative = "two.sided")

```

This result matches the power calculated using G*Power Exact Tests of Proportions, assuming large sample approximation for testing the inequality of proportions with unconstrained margins. 

```{r}
power.prop.test(n = NULL, p1 = 0.28, p2 = 0.52, sig.level = 0.05,
                power = 0.8,
                alternative = "two.sided")
```

The code above calculates the number of subjects required per group. The power.prop.test assumes equal group size, but a simple simulation can be written to match these results and estimate achieved power with different groups sizes.

ADD IN SIMULATION

If margins are constrained (both the split of observations into the two groups and the number of outcomes are fixed), then the power.fisher.test in the statmod package can be used.

```{r}
library(statmod)
set.seed(1234)
power.fisher.test(.28, .52, 25, 25, alpha=0.05, nsim=10000, alternative="two.sided")
```

Though this is an exact result from 10000 simulated grids, the result is consistent with the G*Power result for Fisher's exact test using these fixed group sizes and observed proportions.  